Journey: A Study in Unhelpful Titles (TODO: should we get a new title?)

### 2017-04-27 paper.tex
Swapping huge pages
Other edits in results section

@@

\subsubsection{Memory}
Figure \ref{fig:mmap_mem_usage} shows the growth in memory usage of the system
for the workloads described in section \ref{ss_vm_area_struct}. The
\texttt{cont} workload coincides with the \texttt{control} workload as only the
first \texttt{mmap} operation allocates a \texttt{vm\_area\_struct}. The subsequent
operations are subsumed by the initial \texttt{vm\_area\_struct} as the mapped
regions are adjacent, and hence the memory usage does not increase despite the
large number of \texttt{mmap} operations.  The \texttt{stride} workload allocates a
\texttt{vm\_area\_struct} for each operation due to the separation of
allocations in the virtual address space. As a result, the memory usage
increases significantly as more memory is consumed with each \texttt{mmap} operation.
This is also the case for the \texttt{rand} workload which breaks up a single
\texttt{vm\_area\_struct} for a large area in the virtual address space into
multiple smaller areas, which results in the  allocation of a new structure for
every resize operation. The \texttt{rand} workload initially creates a new
\texttt{vm\_area\_struct} struct for each \texttt{mmap} operation. Eventually, these
allocations form larger contiguous allocations when two or more regions are
mapped next to each other, resulting in a single \texttt{vm\_area\_struct} being
used for that range. As a result, the memory usage begins to flatten out and
even decreases slightly as multiple \texttt{vm\_area\_struct}s are collapsed to
single structures. This is made clear by Figure \ref{fig:vm_area_struct_count}
which shows the growth \texttt{vm\_area\_struct} counts for each workload.
\begin{figure}
    \includegraphics[width=\columnwidth]{figures/mmap_mem_usage}
    \caption{$\Delta_i$ adjusted with control for benchmarks}
    \label{fig:mmap_mem_usage}
\end{figure}

@@

\subsubsection{Memory}
Figure \ref{fig:swap_rss} shows the Resident Set Size of the system for the
experiments described in section \ref{swapping_memory_overhead}. The
\texttt{cont} workloads write to  adjacent 4KB pages in the address space. A
huge page of size 2MB is allocated (TODO: Is it 2MB?) when a new page is touched
initially. Subsequent 4KB pages that are touched are subpages of the 2MB page
allocated initially. As a result, the memory used by the workload is almost
equal to the amount of memory it has written to. The RSS thus increases linearly
until the system begins to swap when the number of free pages falls below a
threshold. The \texttt{rand} workload on the other hand writes to pages that are
not adjacent, and likely not subpages of a single huge page. As a result, a huge
page is allocated with almost every write, and the system starts to exhaust its
memory and begins swapping very quickly. Memory made available by kswapd is
quickly consumed by further writes and hence the memory used by the workload
continues to stay close to its peak value. As expected, the behavior is the same
irrespective of the size of the swap file.

>>>>

(TODO: we need to be consistent with whether we texttt the benchmark names or
not)

@@

\subsubsection{Memory}
Figure \ref{fig:mmap_mem_usage} shows the growth in memory usage of the system
for the workloads described in section \ref{ss_vm_area_struct}.

The
\texttt{cont} workload coincides with the \texttt{control} workload, as only the
first \texttt{mmap} operation allocates a \texttt{vm\_area\_struct}. The subsequent
operations are subsumed by the initial \texttt{vm\_area\_struct} because the mapped
regions are adjacent, and hence the memory usage does not increase despite the
large number of \texttt{mmap} operations.

The \texttt{stride} workload allocates a
\texttt{vm\_area\_struct} for each operation due to the separation of
allocations in the virtual address space. As a result, the memory usage
increases significantly as more memory is consumed with each \texttt{mmap} operation.

This is also the case for the \texttt{frag} workload which breaks up a single
\texttt{vm\_area\_struct} for a large area in the virtual address space into
multiple smaller areas, which results in the allocation of a new structure for
every resize operation.

The \texttt{rand} workload initially creates a new
\texttt{vm\_area\_struct} struct for each \texttt{mmap} operation. Eventually, these
allocations form larger contiguous allocations when two or more regions are
mapped next to each other, resulting in a single \texttt{vm\_area\_struct} being
used for that range. As a result, the memory usage begins to taper off and
even decreases slightly as multiple \texttt{vm\_area\_struct}s are collapsed to
single structures. This is made clear by Figure \ref{fig:vm_area_struct_count}
which shows the growth \texttt{vm\_area\_struct} counts for each workload.

TODO

\begin{figure}
    \includegraphics[width=\columnwidth]{figures/mmap_mem_usage}
    \caption{$\Delta_i$ adjusted with control for benchmarks}
    \label{fig:mmap_mem_usage}
\end{figure}

@@

\subsubsection{Memory}

Figure \ref{fig:swap_rss} shows the RSS of each workload for the
experiments described in section \ref{swapping_memory_overhead}.

The
cont workloads write to  adjacent 4KB pages in the address space. A
huge page of size 2MB is allocated when a new page is touched
initially. Subsequent 4KB pages that are touched are subpages of the 2MB page
allocated initially. As a result, the memory used by the workload is almost
equal to the amount of memory it has written to. The RSS thus increases linearly
until the system begins to swap when the number of free pages falls below a
threshold.

The rand workload, on the other hand, writes to pages that are likely
not adjacent and likely not subpages of a single huge page. As a result, a huge
page is allocated with almost every write, and system memory is exhausted and
begins swapping very quickly. Memory made available by kswapd is
quickly consumed by further allocations. Hence, the memory used by the workload
continues to stay close to its peak value. As expected, the behavior is the same
irrespective of the size of the swapfile.

TODO
